Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-12-24 paper/paper.tex
update

@@

\section{Benchmarks and case studies}
\subsection{Performance benchmarks}
We run a single thread benchmark on CPU Intel(R) Xeon(R) CPU E5-2686 v4 @ 2.30GHz,
and its CUDA version on a GPU Tesla V100.
The results are summarized in Figure~\ref{fig:benchmark}.
The graphs that we use in benchmarks are random 3 regular graphs, a typical type of sparse graphs that has a small tree width $\text{tw}(G) \leq |V|/6$~\cite{Fomin2006}.

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction process, without taking the just in time compiling time into account.
    \red{for each graph size, is there only one graph?}\blue{yes}
    (a) treewidth versus the number of vertices for the benchmarked graphs. 
    Legends are the data types that we used in the computation, one can find the corresponding computational task in \Tbl{tbl:generictypes}.
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for enumeration problems, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    \blue{If we use markers, the data poitns are too dense. I prefer not adding markers, not much people read papers with kindle or on paper.}
    \blue{I used a smaller fontsize, there are still some overlaps. Does it look better (or worse) now?}
    }
    \label{fig:benchmark}
\end{figure}

The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
computing independence polynomial (panel c) and enumerating and sampling independent sets (panel d).
\blue{I cleaned up this paragraph, because one can check the table \Tbl{tbl:generictypes} now.}
\red{Jinguo, should we include maximal independence polynomial and enumeration of all maximal independent sets as well? I think that will make the benchmark more complete.}
\blue{added a new paragraph.}
Among the benchmarks, computing \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions, and hence 
they are faster comparing to non-BLAS methods in the same category.
Most algebras can be computed on GPU, except those requiring dynamic sized structures, i.e. \texttt{PN} and \texttt{SN}.
Fourier transformation based method is the fastest in computing the independence polynomial,
while the finite field (GF$(p)$) approach is the only method that does not incur round-off errors and that can be run on a GPU.
In the benchmark of enumerating the MISs (panel d), one can see the technique to bound the enumeration space improves the performance for more than one order.
Bounding can also reduce the memory usage significantly, without which the largest enumerable size is about $150$ vertices on a 32G memory device.
All these computing times have a strong correlation with the treewidth, 
and the computing times of configuration enumeration also depend on the size of the target configuration space.

The standard algorithm to enumerate all maximal independent sets is the Bron-Kerbosch algorithm~\cite{Bron1973}, which, of course, includes the enumeration of all MISs and can be used to enumerate all independent sets as well.
We also benchmarked our method with the Bron-Kerbosch method to enumerate and count maximal independent sets in \Fig{fig:benchmark-maximal}.
Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
The implementation of Bron-Kerbosch algorithm is from the package LightGraphs~\cite{LightGraphs},
due to the memory limit, this algorithm stops working at size $70$ and above.

@@

\begin{description}
	\item[\href{https://github.com/under-Peter/OMEinsum.jl}{OMEinsum} and \href{https://github.com/Happy-Diode/OMEinsumContractionOrders.jl}{OMEinsumContractionOrders}] are packages providing the support for Einstein's (or tensor network) notation and contraction order optimizations.
    \texttt{OMEinsumContractionOrders} implements state of art algorithms for finding the optimal contraction order for the an tensor network, including the KaHypar+Greedy~\cite{Gray2021, Pan2021} and local transformation based approaches~\cite{Kalachev2021},
	\item[\href{https://github.com/TensorBFS/TropicalNumbers.jl}{TropicalNumbers} and \href{https://github.com/TensorBFS/TropicalGEMM.jl}{TropicalGEMM}] are packages providing tropical number and efficient tropical matrix multiplication,
	\item[\href{https://github.com/JuliaGraphs/LightGraphs.jl}{LightGraphs}] is a package providing graph utilities, like random regular graph generator,
	\item[\href{https://github.com/JuliaMath/Polynomials.jl}{Polynomials}] is a package providing polynomial algebra and polynomial fitting,
	\item[\href{https://github.com/scheinerman/Mods.jl}{Mods} and \href{https://github.com/JuliaMath/Primes.jl}{Primes}] are packages providing finite field algebra and prime number generations.
\end{description}

One can install these packages by opening a Julia REPL, type \colorbox{lightgray}{\texttt{]}} to enter the \texttt{pkg>} mode and type, e.g.
\begin{lstlisting}
pkg> add OMEinsum LightGraphs Mods Primes Polynomials TropicalNumbers OMEinsumContractionOrders
\end{lstlisting}

@@

@@

>>>>

\section{Benchmarks and case studies}
\subsection{Performance benchmarks}
We run a single thread benchmark on CPU Intel(R) Xeon(R) CPU E5-2686 v4 @ 2.30GHz,
and its CUDA version on a GPU Tesla V100.
The results are summarized in Figure~\ref{fig:benchmark}.
The graphs that we use in benchmarks are random three regular graphs,
 a typical type of sparse graphs that has a small tree width asymptotically smaller than $|V|/6$~\cite{Fomin2006}.

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction, without taking the contraction order finding and just in time compiling time into account.
    \red{for each graph size, is there only one graph?}\blue{yes}
    (a) tree width versus the number of vertices for the benchmarked graphs. 
    Legends are the data types that we used in the computation, one can find the corresponding computed property in \Tbl{tbl:generictypes}.
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for configuration enumeration, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    \blue{If we use markers, the data poitns are too dense. I prefer not adding markers, not much people read papers with kindle or on paper.}
    \blue{I used a smaller fontsize, there are still some overlaps. Does it look better (or worse) now?}
    }
    \label{fig:benchmark}
\end{figure}

Figure (a) shows the time and space complexity obtained with tensor network optimization algorithm in Ref.~\cite{Kalachev2021}, where the space complexity is the same as the tree width of the problem graph.
One can see all the computing times in figure (b), (c) and (d) have a strong correlation with the treewidth.
%The computations shown in \Fig{fig:benchmark} include independent set counting, finding maximum independent set size (panel b),
%computing independence polynomial (panel c) and find one independent set and enumerating independent sets (panel d).
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions, hence 
are much faster comparing to non-BLAS methods in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough so that the actual computing time is comparable or larger than the launching time of CUDA kernels.
Most algebras can be computed on GPU, except those requiring dynamic sized structures, i.e. \texttt{PN} and \texttt{SN}.
In figure (c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
however, it maybe suffer from the round off errors. The finite field (GF$(p)$) approach is the only method that does not incur round-off errors and can run on a GPU.
In figure (d), one can see the technique to bound the enumeration space improves the performance for more than one order in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a 32G memory device.
%and the computing times of configuration enumeration also depend on the size of the target configuration space.

@@

@@

\begin{description}
	\item[\href{https://github.com/under-Peter/OMEinsum.jl}{OMEinsum} and \href{https://github.com/Happy-Diode/OMEinsumContractionOrders.jl}{OMEinsumContractionOrders}] are packages providing the support for Einstein's (or tensor network) notation and contraction order optimizations.
    \texttt{OMEinsumContractionOrders} implements state of art algorithms for finding the optimal contraction order for the an tensor network, including the KaHypar+Greedy~\cite{Gray2021, Pan2021} and local transformation based approaches~\cite{Kalachev2021},
	\item[\href{https://github.com/TensorBFS/TropicalNumbers.jl}{TropicalNumbers} and \href{https://github.com/TensorBFS/TropicalGEMM.jl}{TropicalGEMM}] are packages providing tropical number and efficient tropical matrix multiplication,
	\item[\href{https://github.com/JuliaGraphs/Graphs.jl}{Graphs}] is a package providing graph utilities, like random regular graph generator,
	\item[\href{https://github.com/JuliaMath/Polynomials.jl}{Polynomials}] is a package providing polynomial algebra and polynomial fitting,
	\item[\href{https://github.com/scheinerman/Mods.jl}{Mods} and \href{https://github.com/JuliaMath/Primes.jl}{Primes}] are packages providing finite field algebra and prime number generations.
\end{description}

One can install these packages by opening a Julia REPL, type \colorbox{lightgray}{\texttt{]}} to enter the \texttt{pkg>} mode and type, e.g.
\begin{lstlisting}
pkg> add OMEinsum Graphs Mods Primes Polynomials TropicalNumbers OMEinsumContractionOrders
\end{lstlisting}

@@

Here is a short introduction to the functionalities in this libary.

\lstinputlisting[breaklines]{../democode/tutorial.jl}

@@

We show the benchmark of computing the maximal independent set properties in \Fig{fig:benchmark-maximal},
including a comparison to the Bron-Kerbosch algorithm from the package Graphs~\cite{Graphs}.
Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
Due to the memory limit, this Bron-Kerbosch algorithm stops working at size $70$ and above.
